---
title: "hw6"
author: "Akhilesh Mahajan"
date: "April 1, 2018"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(rvest)
library(tidyverse)
```

1. Taking advantage of the rvest package, turn the table at http://www.nature.com/articles/ng.3097/tables/3 into an R data frame

Ans:
```{r}
#Used selector gadget for finding the nodes.
genes <- "http://www.nature.com/articles/ng.3097/tables/3" %>% 
  read_html() %>%
  html_nodes("#content .mq1200-padded") %>%
  html_nodes("table") %>%
  html_table(fill = TRUE) %>% 
  as.data.frame()
```

a. Be sure to delete the rows "Genes with" and "New genes without" (and don't do it by using the row number).

Ans:
```{r}
# Since these particular rows have same values over all the columns, we filter these rows by randomly selecting any column
genes <- genes %>% 
  filter(!grepl("Genes with *", Gene)) %>%
  filter(!grepl("New genes without *", Gene))
```


b. Be sure to convert the p value column to numbers

Ans:P value cleaning function
```{r}
p.column.to.number <- function(x) {
  #We split on space
  temp <- str_split(x, " ") %>% unlist(.)
  base <- temp[1]
  
  if(!is.na(temp[3])) {
    #There is explicit exponent present
    #We extract all the digits present in the exponent
    exponent <- str_extract(temp[3], "[0-9]{1,}$")
    as.double(paste0(base, "e", "-", exponent))
  } else {
    #There is no explicit exponent present
    as.double(paste0(base))
  }
}
```

```{r}
genes.p.list <- genes %>% select(Prioritization.P.value) %>% apply(1, FUN = function(x) p.column.to.number(x))
genes <- genes %>% mutate(Prioritization.P.value = genes.p.list)
```

c. The last column is shown in 3 rows in the journal, but most likely as one string in your table. Use regular expressions to insert semicolons (";") between each of the original lines. For example,
    "PI3K cascade (REACTOME, P = 6.2 × 10−13); Chronic myeloid leukemia (KEGG, P = 1.6 × 10−12); Response to fibroblast growth factor stimulus (GO, P = 5.4 × 10−11)"
    

Ans:We observer that whenever we have ')' closing round brackets, we need a semicolon after that.

```{r}
semicolon.adder <- function(x) {
  x %>% str_replace_all("\\)", "\\);")
}

gene.set.list <- genes %>% select(Top.ranking.reconstituted.gene.sets) %>% apply(1, FUN = function(x) semicolon.adder(x))
genes <- genes %>% mutate(Top.ranking.reconstituted.gene.sets = gene.set.list)
```


d. Show the table in your RMarkdown file by using the function `knitr::kable()`

Ans:We see that the data frame **genes** loses it's value in column Prioritization.P.value. Hence we add digits = 20 argument.
```{r}
knitr::kable(genes, digits = 20)
```