---
title: "hw4"
author: "Akhilesh Mahajan"
date: "February 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(babynames)
library(Lahman)
library(lubridate)
library(stringr)
library(curl)
library(gapminder)
```


## Question and Answers:
1) 

a) Create a new function that, given an `lm` object, returns the n residuals with the largest absolute values (but return the residuals, not the absolute value of the residuals), where the default value for n is 5. The function should give a clear error message if n is larger than the number of residuals. Demonstrate that your function works by applying it to
        mtcars.lm <- lm(mpg ~ disp, data = mtcars)
    first with no argument for n, then with n = 6, and then with n = 40.

Ans: 
```{r}
mtcars.lm <- lm(mpg ~ disp, data = mtcars)

residuals.get <- function(objec, n=5){
  resid <- objec$residuals
  sort_resid <- resid[order(abs(resid))]
  try(if(n > nrow(mtcars)) stop("n is larger than number of residuals"))
  if(n<=nrow(mtcars)){
    tail(sort_resid,n)
  }
}

residuals.get(mtcars.lm)
residuals.get(mtcars.lm, n=6)
residuals.get(mtcars.lm, n=40)
```




2) Read the file "height.txt", in the week 4 dropbox folder, and use regular expressions to clean the height variable and convert it into a numeric column representing height in inches. Make a table showing the number of non-missing values of height for men and women. Finally, make a plot for height, showing the density for men and for women on the same plot.

Ans:

```{r}
clean_height <- function(x) {
#  ifelse(str_detect(x, "[2-9]{1}'[0-9]{1}|[2-9]{1}'[1]{1}[0-1]{1}"), micro_function1(x), x)
  
  # correct data
  if (str_detect(x, "^[2-9]{1}[0-9]{1}$|^[1]{1}[0]{1}[0-9]{1}$")) {
    x
  }
  # cases like 5'7
  else if (str_detect(x, "^[2-9]{1}'[1]{1}[0-1]{1}|[2-9]{1}'[0-9]{1}$")) {
    temp <- str_split(x, "'") %>% unlist(.)
    inches <- 12*as.numeric(temp[1]) + as.numeric(temp[2])
  }
  # cases like 5,7
  else if (str_detect(x, "^[2-9]{1},[1]{1}[0-1]{1}$|^[2-9]{1},[0-9]{1}$")) {
    temp <- str_split(x, ",") %>% unlist(.)
    inches <- 12*as.numeric(temp[1]) + as.numeric(temp[2])
  }
  # cases like 178 - cm to inches
  else if (str_detect(x, "^[1]{1}[1-9]{1}[0-9]{1}|[2]{1}[0-9]{1}[0-9]{1}c?m?$")) {
    temp <- str_split(x, "cm") %>% unlist(.)
    inches <- round(as.numeric(temp[1]) * 0.393)
  }
  # cases like 64.234 to 64
  else if (str_detect(x, "^[2-9]{1}[0-9]{1}.|^[1]{1}[0]{1}[0-9]{1}.")) {
    temp <- str_split(x, "\\.") %>% unlist(.)
    inches <- temp[1]
  }
  # cases like 5.4
  else if (str_detect(x, "^[2-9]{1}.[0-9]{1}$")) {
    temp <- str_split(x, "\\.") %>% unlist(.)
    inches <- as.numeric(temp[1])*12 + as.numeric(temp[2])*1.2
    inches <- round(inches)
  }
  # cases like 5.25
  else if (str_detect(x, "^[2-9]{1}.[0-9]{2}$")) {
    temp <- str_split(x, "\\.") %>% unlist(.)
    inches <- as.numeric(temp[1])*12 + as.numeric(temp[2])*0.12
    inches <- round(inches)
  }
  # cases like 5
  else if (str_detect(x, "^[2-9]{1}$")) {
    inches <- as.numeric(x)*12
  }
  # cases like 5' 11
  else if (str_detect(x, "^[2-9]{1}' [1]{1}[0-1]{1}$|^[2-9]{1}' [0-9]{1}$")) {
    temp <- str_split(x, "' ") %>% unlist(.)
    inches <- 12*as.numeric(temp[1]) + as.numeric(temp[2])
  }
  # cases like 6 04
  else if (str_detect(x, "^[2-9]{1} [1]{1}[0-1]{1}$|^[2-9]{1} [0-9]{1}$")) {
    temp <- str_split(x, " ") %>% unlist(.)
    inches <- 12*as.numeric(temp[1]) + as.numeric(temp[2])
  }
  else {
    NA
  }
}
```



```{r}
height.data <- read_tsv("~/Downloads/data_wrangling/week4/height.txt")
height.data.dirty <- height.data
#height.data.clean <- height.data %>% filter(str_detect(height, "^[2-9]{1}[0-9]{1}$|^[1]{1}[0]{1}[0-9]{1}$"))
#height.data.dirty <- height.data %>% anti_join(height.data.clean)
height.data.dirty <- height.data.dirty %>% mutate("height.clean" = str_replace_all(height, "^[^[:alnum:]]{0,}", ""))
height.data.dirty <- height.data.dirty %>% mutate("height.clean" = str_replace_all(height.clean, "[^[:alnum:]]{0,}$", ""))

height.cleaned.list <- height.data.dirty %>% select(height.clean) %>% apply(1, FUN = function(x) clean_height(x)) 
height.data.dirty <- height.data.dirty %>% mutate(height.cleaned = as.numeric(height.cleaned.list))
```
Table for non-missing height for men and women count:
```{r}
na.omit(height.data.dirty) %>% group_by(sex) %>% summarize(count = n())
```

Plot for height:
```{r}
final.data <- na.omit(height.data.dirty)
final.data %>% ggplot(aes(x = height.cleaned)) + geom_density(colour = "black", fill = "#56B4E9") + facet_wrap(~ sex) + scale_x_continuous(name = "Height in inches", limits = c(0, 90)) + ggtitle("Density plot of heights of males and females")
```

0) Find all the entries that are in inches. Normal height is from 20 inches to 109 inches. Use this normal height to find dirty heights.
1) Remove "" '', any special characters from the start and end. Remove [[:space:]] and [[:punct:]].
2) Now convert 5'11, 5.11, 5' 11, 5.25 to inches. This conversion is from feet to inches. Range is from 2'0 to 9'11.
3) special cases include: 
66, 6, "5' 4""", 5.3, 165cm, 511, 2, 5'7, >9000, 5 feet and 8.11 inches, 5.25, 5'11, 11111, 5'9'', 150, 5,3, 6', 5' 10, Five foot eight inches, 5.75, 5'10'',
0.7, 184, "5' 7.78""", 12, yyy, 5 feet 7inches, 1, 1.7, "69""", 5ft 9 inches, 5 ft 9 inches, 5'11'', 708,661, 649,606, 6 04, 0, 170 cm, 7,283,465




3) Split the gapfinder data by country and use map() to calculate, by country, the r^2 for the linear model lifeExp ~ log10(gdpPercap). Using ggplot2, make a set of boxplots of r^2 by continent.


Ans:
```{r}
gapminder.split <- gapminder %>% split(.$country)
gapminder.lm <- gapminder.split %>% map(~ lm(lifeExp ~ log10(gdpPercap), weights = pop, data = .))
r.squared.list <- map(gapminder.lm, ~ summary(.x)$r.squared)
y <- stack(r.squared.list)
r.list.joined <- inner_join(y, gapminder, by = c("ind" = "country")) %>% select(values, ind, continent)
r.list.joined <- unique(r.list.joined)
r.list.joined %>% ggplot(aes(x = continent, y = values, fill = continent)) + geom_boxplot() + ggtitle("Boxplot for the continent and their corresponding countries' r square values")
```