---
title: "hw8"
author: "Akhilesh Mahajan"
date: "April 7, 2018"
output: html_document
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(DBI)
library(dbplyr)
library(Modeler)
library(pool)
library(RMySQL)
library(gapminder)
library(broom)
```


1) The following code will set up a connection to a demo database. You will probably need to install the `pool` package.

Ans:

```{r}
my_db <- dbPool(
  RMySQL::MySQL(), 
  dbname = "shinydemo",
  host = "shiny-demo.csa7qlmguqrf.us-east-1.rds.amazonaws.com",
  username = "guest",
  password = "guest"
)
```

This database has three tables, City, Country, and CountryLanguage.

```{r}
city <- collect(tbl(my_db, "City"))
country <- collect(tbl(my_db, "Country"))
countryLanguage <- collect(tbl(my_db, "CountryLanguage"))
```

a) By joining the City and Country tables on the code variable (slightly different names in each table), create a new table containing only the name of the city,  the continent, the region, and the population of the city (not the population of the country). 

```{r}
city.country.joined <- city %>% 
  inner_join(country, by = c("CountryCode" = "Code")) %>%
  select(Name.x, Continent, Region, Population.x)
```

b) Restricting to North America, make boxplots of population by region.

```{r}
city.country.joined %>% 
  filter(Continent == "North America") %>% 
  ggplot(aes(x = Region, y = Population.x)) + 
  geom_boxplot() + ggtitle("Population boxplot of North America by region") +
  scale_y_continuous("Population")
```

c) Taking advantage of the `nest()` function, show the 5 largest cities in each region in North America.

The following cities are the largest cities by population for each region of North America.
```{r}
city.country.joined.nested <- city.country.joined %>%
  filter(Continent == "North America") %>% 
  group_by(Region) %>% 
  nest()

city.country.joined.nested$data %>% 
  map(~ .x %>% 
        arrange(desc(Population.x)) %>%
        head(n = 5) %>% 
        select(Name.x)) %>%
  set_names(city.country.joined.nested$Region)
```

2, There is a built-in dataframe called "ChickWeight"

a) Using `nest()`, construct a 50 x 3 dataframe with columns corresponding to the chick id, the chick diet, and dataframes of each chick's data

```{r}
ChickWeight.nested <- ChickWeight %>% 
  group_by(Chick, Diet) %>% 
  nest()
```

b) Create a new column consisting of an `lm` model for each chick, where the regression is `weight ~ Time`


```{r}
ChickWeight.nested.lm <- ChickWeight.nested %>% 
  mutate(lm = map(data, ~ lm(data = ., weight ~ Time)))
```

c) Add columns to the dataframe giving the slope and intercept for each chick (the `broom` package might be helpful, but it's not the only way).


```{r}
ChickWeight.nested.lm.slope.intercept <- ChickWeight.nested.lm %>% 
  mutate(intercept = map(lm, ~ as.numeric(.x$coefficients[1])) %>% unlist(),
         slope = map(lm, ~ as.numeric(.x$coefficients[2])) %>% unlist())
```

d) Make a scatterplot of slope (vertical axis) and intercept (horizontal axis) colored by Diet


We see that when slope is very high, the intercept is low. This graph itself somewhat follows linear curve.
```{r}
ChickWeight.nested.lm.slope.intercept %>%
  ggplot(aes(x = intercept, y = slope, color = Diet)) + 
  geom_point() + geom_line()
  ggtitle("Scatterplot of Slope and Intercept")
```