---
title: "hw5"
author: "Akhilesh Mahajan"
date: "February 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(lubridate)
library(stringr)
library(gapminder)
library(gutenbergr)
library(tidytext)
library(readxl)
library(purrr)
install.packages("data.table")
library(data.table)
library(ggrepel)
```


## Question and Answers:
1) 
a) Clean as many of the height answers as you can; do not replace values with hard-coded numbers (i.e., don't type height[29] <- 62). Do so by creating a new variable for height; don't replace values in the original variable.

Ans:
```{r}
clean_height <- function(x) {
#  ifelse(str_detect(x, "[2-9]{1}'[0-9]{1}|[2-9]{1}'[1]{1}[0-1]{1}"), micro_function1(x), x)
  
  # correct data
  if (str_detect(x, "^[2-9]{1}[0-9]{1}$|^[1]{1}[0]{1}[0-9]{1}$")) {
    x
  }
  # cases like 5'7
  else if (str_detect(x, "^[2-9]{1}'[1]{1}[0-1]{1}|[2-9]{1}'[0-9]{1}$")) {
    temp <- str_split(x, "'") %>% unlist(.)
    inches <- 12*as.numeric(temp[1]) + as.numeric(temp[2])
  }
  # cases like 5,7
  else if (str_detect(x, "^[2-9]{1},[1]{1}[0-1]{1}$|^[2-9]{1},[0-9]{1}$")) {
    temp <- str_split(x, ",") %>% unlist(.)
    inches <- 12*as.numeric(temp[1]) + as.numeric(temp[2])
  }
  # cases like 178 - cm to inches
  else if (str_detect(x, "^[1]{1}[1-9]{1}[0-9]{1}|[2]{1}[0-9]{1}[0-9]{1}c?m?$")) {
    temp <- str_split(x, "cm") %>% unlist(.)
    inches <- round(as.numeric(temp[1]) * 0.393)
  }
  # cases like 64.234 to 64
  else if (str_detect(x, "^[2-9]{1}[0-9]{1}.|^[1]{1}[0]{1}[0-9]{1}.")) {
    temp <- str_split(x, "\\.") %>% unlist(.)
    inches <- temp[1]
  }
  # cases like 5.4
  else if (str_detect(x, "^[2-9]{1}.[0-9]{1}$")) {
    temp <- str_split(x, "\\.") %>% unlist(.)
    inches <- as.numeric(temp[1])*12 + as.numeric(temp[2])*1.2
    inches <- round(inches)
  }
  # cases like 5.25
  else if (str_detect(x, "^[2-9]{1}.[0-9]{2}$")) {
    temp <- str_split(x, "\\.") %>% unlist(.)
    inches <- as.numeric(temp[1])*12 + as.numeric(temp[2])*0.12
    inches <- round(inches)
  }
  # cases like 5
  else if (str_detect(x, "^[2-9]{1}$")) {
    inches <- as.numeric(x)*12
  }
  # cases like 5' 11
  else if (str_detect(x, "^[2-9]{1}' [1]{1}[0-1]{1}$|^[2-9]{1}' [0-9]{1}$")) {
    temp <- str_split(x, "' ") %>% unlist(.)
    inches <- 12*as.numeric(temp[1]) + as.numeric(temp[2])
  }
  # cases like 6 04
  else if (str_detect(x, "^[2-9]{1} [1]{1}[0-1]{1}$|^[2-9]{1} [0-9]{1}$")) {
    temp <- str_split(x, " ") %>% unlist(.)
    inches <- 12*as.numeric(temp[1]) + as.numeric(temp[2])
  }
  else {
    NA
  }
}
```



```{r}
height.data <- read_tsv("~/Downloads/data_wrangling/week4/height.txt")
height.data.dirty <- height.data
#height.data.clean <- height.data %>% filter(str_detect(height, "^[2-9]{1}[0-9]{1}$|^[1]{1}[0]{1}[0-9]{1}$"))
#height.data.dirty <- height.data %>% anti_join(height.data.clean)
height.data.dirty <- height.data.dirty %>% mutate("height.clean" = str_replace_all(height, "^[^[:alnum:]]{0,}", ""))
height.data.dirty <- height.data.dirty %>% mutate("height.clean" = str_replace_all(height.clean, "[^[:alnum:]]{0,}$", ""))

height.cleaned.list <- height.data.dirty %>% select(height.clean) %>% apply(1, FUN = function(x) clean_height(x)) 
height.data.dirty <- height.data.dirty %>% mutate(height.cleaned = as.numeric(height.cleaned.list))
```

b) Provide a table of the number of missing values for the new variable by sex.
Ans:
```{r}
height.data.dirty %>% filter(is.na(height.cleaned)) %>% group_by(sex) %>% summarize(count = n())
```

c) Print all of the original values for height for which your methods could not provide a clean value.
Ans:
```{r}
height.data.dirty %>% filter(is.na(height.cleaned)) %>% select(height) %>% print(n = Inf)
```

d) Plot, in a single plot, the density of height for men and women.
Plot for height:
```{r}
final.data <- na.omit(height.data.dirty)
final.data %>% ggplot(aes(x = height.cleaned)) + geom_density(colour = "black", fill = "#56B4E9") + facet_wrap(~ sex) + scale_x_continuous(name = "Height in inches", limits = c(0, 90)) + ggtitle("Density plot of heights of males and females")
```


2) 

a) Download the texts of Wuthering Heights and Jane Eyre, using the gutenbergr package.

Ans: 
```{r}
bronte_books <- gutenberg_download(gutenberg_id = c(768, 1260), meta_fields = "title")
```

b) Find the 10 most common non-stop-words in Jane Eyre.

Ans:
```{r}
data("stop_words")
bronte_books %>% filter(title == "Jane Eyre: An Autobiography") %>% 
  mutate(linenumber = row_number(), chapter = cumsum(str_detect(text, "(?i)^chapter [0-9ivxlc]"))) %>% 
  unnest_tokens(words, text) %>% 
  anti_join(stop_words, by = c("words" = "word")) %>% 
  group_by(words) %>% 
  summarize(count = n()) %>% 
  arrange(-count) %>% 
  head(n = 10)
```

c) Find the 10 most common bigrams in Jane Eyre that do not include stop words.

Ans:
```{r}
bronte_books %>% filter(title == "Jane Eyre: An Autobiography") %>% 
  mutate(linenumber = row_number(), chapter = cumsum(str_detect(text, "(?i)^chapter [0-9ivxlc]"))) %>% 
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>% 
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>% 
  filter(!word2 %in% stop_words$word) %>% 
  unite(bigram, word1, word2, sep = " ") %>% 
  group_by(bigram) %>% 
  summarize(count = n()) %>% 
  arrange(-count) %>% 
  head(n = 10)
```

d) Plot the sentiment for the two books.

Ans:
```{r}
bronte_books %>% group_by(title) %>%
  mutate(linenumber = row_number(), chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text) %>% 
  filter(word != "miss") %>%
  inner_join(get_sentiments("bing")) %>%
  count(title, index = linenumber %/% 80, sentiment) %>% 
  spread(sentiment, n, fill = 0) %>% 
  mutate(sentiment = positive - negative) %>% 
  ggplot(aes(index, sentiment, fill = title)) + geom_bar(alpha = 0.8, stat = "identity", show.legend = FALSE) + facet_wrap(~title, ncol = 3, scales = "free_x") + theme(legend.position = "none") + geom_smooth(span = 0.15)
```

3) The code appended to the end of this file will download some data collected by www.gapminder.org/data and then read it into a list of wide data frames (notice the use of map()). Be sure to include the code in your Rmarkdown files so that it is self-contained.


```{r}
urls <- c("https://docs.google.com/spreadsheet/pub?key=phAwcNAVuyj0NpF2PTov2Cw&output=xlsx",
          "https://docs.google.com/spreadsheet/pub?key=phAwcNAVuyj2tPLxKvvnNPA&output=xlsx",
          "https://docs.google.com/spreadsheet/pub?key=phAwcNAVuyj0TAlJeCEzcGQ&output=xlsx",
          "https://docs.google.com/spreadsheet/pub?key=phAwcNAVuyj0XOoBL_n5tAQ&output=xlsx",
          "https://docs.google.com/spreadsheet/pub?key=0AkBd6lyS3EmpdHo5S0J6ekhVOF9QaVhod05QSGV4T3c&output=xlsx",
          "https://docs.google.com/spreadsheet/pub?key=pyj6tScZqmEcVezxiMlWaRw&output=xlsx",
          "https://docs.google.com/spreadsheet/pub?key=pyj6tScZqmEeL79qOoKtofQ&output=xlsx",
          "https://docs.google.com/spreadsheet/pub?key=pyj6tScZqmEcjeKHnZq6RIg&output=xlsx"
          )

measures <- c("infant_mortality", "life_expectancy",
                 "fertility","population","gdp", "maternal_mortality", "health_spending", "gini")



for (i in 1:length(urls)){
  if (!file.exists(paste0(measures[i], ".xlsx")))
  download.file(urls[i], destfile = paste0(measures[i], ".xlsx"), method = "auto")
}

gap_raw <- map(paste0(measures, ".xlsx"), read_excel, na = c("", "NA", "-", "."))
names(gap_raw) <- measures
```

a) Make a new list, where each list element is a tall data frame. Try to use functions from the purrr package when you do so.

Ans:
```{r}
gap_raw_tall <- gap_raw %>% 
  map(~ gather(.x[, -1], key = "year", value = "count")) %>%
  map2(.x = ., .y = gap_raw, .f = ~ data.frame(.x, .y[, 1])) %>%
  map(~ .x[complete.cases(.x), ]) %>%
  map(~ setnames(.x, old = 3, new = "country"))
```

b) Join the tall data frames together to get a single data frame much like the gapminder data frame in the package gapminder. *Do not call your new data frame "gapminder"*---that would be too confusing.

Ans:
```{r}
a <- data.frame(gap_raw_tall$infant_mortality)
b <- data.frame(gap_raw_tall$life_expectancy)
c <- data.frame(gap_raw_tall$fertility)
d <- data.frame(gap_raw_tall$population)
e <- data.frame(gap_raw_tall$gdp)
f <- data.frame(gap_raw_tall$maternal_mortality)
g <- data.frame(gap_raw_tall$health_spending)
h <- data.frame(gap_raw_tall$gini)
gapminder.new <- Reduce(function(x, y) merge(x, y, by = c("year", "country"), all = TRUE), list(a, b, c, d, e, f, g, h))
setnames(gapminder.new, old = 3:10, new = c("infant_mortality", "life_expectancy", "fertility", "population", "gdp", "maternal_mortality", "health_spending", "gini"))
```



c) Make one plot showing health_spending over time since 1945 for the 12 richest countries (as measured by gdp per capita at the latest available year). Be sure to make the plot look finished (use a title, change axes labels if necessary, etc.)
Ans:
```{r}
# Checking if any of the year values are null; further program depends on this.
# No null values in year.
gapminder.new %>% select(year) %>% filter(is.na(.)) 

gapminder.new %>%
  filter(year > 1944) %>%
  select(year, country, health_spending) %>% 
  na.omit() %>%
  inner_join(gapminder.new %>%
               select(country, year, gdp) %>%
               arrange(desc(year)) %>%
               na.omit() %>%
               subset(!duplicated(country)) %>%
               arrange(desc(gdp)) %>%
               head(n = 12), by = c("country" = "country")) %>%
  select(1:3) %>%
  ggplot(aes(x = year.x, y = health_spending, color = country)) +
  geom_point() + ggtitle("Health Spending for 12 Richest countries") + 
  theme(axis.text.x = element_text(angle = 80, size = 10)) + 
  scale_x_discrete(name = "Year") + scale_y_continuous(name = "Health Spending") + 
  geom_line(aes(group = country))
```

d) For the same countries and years as in (c), plot maternal mortality (y-axis) against health spending (x-axis). Make the plot look finished.
Ans:
```{r}
gapminder.new %>%
  filter(year > 1944) %>%
  select(year, country, health_spending, maternal_mortality) %>% 
  na.omit() %>%
  inner_join(gapminder.new %>%
               select(country, year, gdp) %>%
               arrange(desc(year)) %>%
               na.omit() %>%
               subset(!duplicated(country)) %>%
               arrange(desc(gdp)) %>%
               head(n = 12), by = c("country" = "country")) %>%
  select(1:4) %>%
  ggplot(aes(x = health_spending, y = maternal_mortality, color = country)) + 
  ggtitle("Health Spending vs Maternal Mortality for 12 richest countries") + 
  scale_x_continuous("Health Spending") + scale_y_continuous("Maternal Mortality") +
  geom_line() + geom_point()
```

e) Using data from one recent year, divide the countries into quintiles based on gdp per capita and make separate plots for each quintile showing
maternal_mortality plotted against the gini index.
Ans:
```{r}
gdp.sorted.countries <- gapminder.new %>% select(year, country, gdp) %>% na.omit() %>% arrange(desc(gdp, year)) %>% subset(!duplicated(country))
```